Effect of quality-adjusted production function inputs

### 2015-05-29 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Calculating dyn lambda for CES.

<<Formulas>>=
# Set up the formulas for each type of analysis.
# These formulas contain energy. 
# We can use keep_summands to drop the energy term (3rd summand on the rhs)
# later if needed.
formulas <- 
  list( 
    PT = c(unadjusted = build_formula("iGDP", list("iKstkS.L", "iL", "iXpMP", "iYear")),
             adjusted = build_formula("iGDP", list("iKservS.L", "ihLest", "iUMP", "iYear"))),
    UK = c(unadjusted = build_formula("iGDP", list("iKstkO.WwithRD", "iL", "iXp", "iYear")),
             adjusted = build_formula("iGDP", list("iKservO.WwithRD", "ihLest", "iU", "iYear")))
  )
countries <- names(formulas)
@

@@

<<Models>>=
flavors <- c("unadjusted", "adjusted")
nests <- list(kle=c(1,2,3), lek=c(2,3,1), ekl=c(3,1,2))
model.types <- c("CD")
# model.types <- c("CD", "CES")

fileName <- "ESEE2015Models.Rdata"
if (!rerunModels){
  # load from the file on disk
  load(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.

@@

dframes <- list()
for (Coun in countries){
  for (flav in flavors){
    forms <- list(noE = keep_summands(formulas[[Coun]][[flav]], -3), withE = formulas[[Coun]][[flav]])
    for (form in forms){
      denom <- summands(form, n=length(summands(form))) # time variable
      nums <- all.vars(form) [1:(length(all.vars(form))-1)] # y, k, l, and e (if present)
      df <- historicalData[[Coun]][c(denom, nums)] # columns are: iYear, y, k, l, and e (if present)
      # Calculate names for the columns of finite difference derivatives
      derivNames <- lapply(1:length(nums), function(i){paste0("d", nums[[i]], "/", "dt")}) %>% unlist
      derivCols <- lapply(1:length(nums), function(i){
        # Calculate finite difference derivative columns
        ediff(df[[i+1]], pad=pad, lag=lag) / ediff(df[[1]], pad=pad, lag=lag)
      }) %>% do.call(cbind, .) %>% as.data.frame
      names(derivCols) <- derivNames
      df <- cbind(derivCols, historicalData[[Coun]][ , c(nums, "Year")])
      e <- if (length(summands(form))==4) "withE" else "noE"
      df$flavor <- if (flav == "unadjusted") "Unadjusted" else "Quality-adjusted"
      df$energy <- if (e == "noE") "Without energy" else "With energy"
      # Add a column for 1/y * dy/dt
#       growth.col.name <- paste0("1/", nums[[1]], "*", derivNames[[1]])
      df[["dydtOvery"]] <- 1 / df[[nums[[1]]]] * df[[derivNames[[1]]]]
      for (m in model.types){
        origCoeffs <- filter(AllCoeffs, model==m & energy==e & Country==Coun & flavor==flav)
        df <- cbind(df, origCoeffs)
        if (m == "CD"){
          lambda_fd <- lapply(1:length(derivNames), function(i){
            # derivative term divided by its numerator.
            # derivative term is in the ith column of df
            # numerator is in i + length(derivNames) th column of df
            out <- df[[i]] / df[[i+length(derivNames)]]
            if (i > 1){
              # Need to make it negative and multiply by the corresponding alpha term.
              # Typical setup is that columns 2 is capital, and we need alpha_1 for capital.
              # So, subtract 1 from i.
              out <- - out * df[[paste0("alpha_", i-1)]]
            }
            return(out)
          }) %>% as.data.frame %>% rowSums
          df$lambda_fd <- lambda_fd
          avg.lambda_fd <- mean(df$lambda_fd, na.rm = TRUE)
          # Fraction of growth explained by Solow Residual
          df$solowFrac <- df$lambda_fd / (df[[1]] / df[[length(derivNames) + 1]])
          avg.solowFrac <- mean(df$solowFrac, na.rm = TRUE)
        } else if (m == "CES"){
          df <- data.frame()
          if (e == "noE"){
            # kl nesting
#             df$lambda_fd <- df[[1]] / 
          } else if (e == "withE"){

          } else {
            stop (paste("Unknown e:", e))
          }

        } else {
          stop(paste("Unknown model", model))
        }
        # Add to the list of data frames
        dframes[[length(dframes)+1]] <- df
      }
    }
  }      
}
temp_data <- do.call(plyr::rbind.fill, dframes)

>>>>

<<Formulas>>=
# Set up the formulas for each type of analysis.
# These formulas contain energy. 
# We can use keep_summands to drop the energy term (3rd summand on the rhs)
# later if needed.
formulas <- 
  list( 
    PT = c(unadjusted = build_formula("iGDP", list("iKstkS.L", "iL", "iXpMP", "iYear")),
             adjusted = build_formula("iGDP", list("iKservS.L", "ihLest", "iUMP", "iYear"))),
    UK = c(unadjusted = build_formula("iGDP", list("iKstkO.WwithRD", "iL", "iXp", "iYear")),
             adjusted = build_formula("iGDP", list("iKservO.WwithRD", "ihLest", "iU", "iYear")))
  )
countries <- names(formulas)
flavors <- names(formulas$PT)
@

@@

<<Models>>=
fileName <- "ESEE2015Models.Rdata"
if (!rerunModels){
  # load from the file on disk
  load(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.

@@

dframes <- list() # Fill with data frames for each combination of country, flavor, energy, model, and nest
for (Coun in countries){
  for (flav in flavors){

    forms <- list(noE = keep_summands(formulas[[Coun]][[flav]], -3), withE = formulas[[Coun]][[flav]])
    for (form in forms){
      denom <- summands(form, n=length(summands(form))) # time variable
      nums <- all.vars(form) [1:(length(all.vars(form))-1)] # y, k, l, and e (if present)
      df <- historicalData[[Coun]][c(denom, nums)] # columns are: iYear, y, k, l, and e (if present)
      # Calculate names for the columns of finite difference derivatives
      derivNames <- lapply(1:length(nums), function(i){paste0("d", nums[[i]], "/", "dt")}) %>% unlist
      derivCols <- lapply(1:length(nums), function(i){
        # Calculate finite difference derivative columns
        ediff(df[[i+1]], pad=pad, lag=lag) / ediff(df[[1]], pad=pad, lag=lag)
      }) %>% do.call(cbind, .) %>% as.data.frame
      names(derivCols) <- derivNames
      df <- cbind(derivCols, historicalData[[Coun]][ , c(nums, "Year")])
      e <- if (length(summands(form))==4) "withE" else "noE"
      df$flavor <- if (flav == "unadjusted") "Unadjusted" else "Quality-adjusted"
      df$energy <- if (e == "noE") "Without energy" else "With energy"
      # Add a column for 1/y * dy/dt
      df[["dydtOvery"]] <- 1 / df[[nums[[1]]]] * df[[derivNames[[1]]]]

#       model.types <- c(CD="CD")
      model.types <- c(CD="CD", CES="CES")
      for (m in model.types){
        if (m == "CD"){
          origCoeffs <- filter(AllCoeffs, model==m & energy==e & Country==Coun & flavor==flav)
          df.CD <- cbind(df, origCoeffs)
          df.CD$nest <- NA
          lambda_fd <- 
            df.CD[["dydtOvery"]] - 
            df.CD[["alpha_1"]] / df.CD[[nums[[2]]]] * df.CD[[derivNames[[2]]]] - 
            df.CD[["alpha_2"]] / df.CD[[nums[[3]]]] * df.CD[[derivNames[[3]]]]
          if (e == "withE"){
            # Subtract the energy term also
            lambda_fd <- lambda_fd - 
              df.CD[["alpha_3"]] / df.CD[[nums[[4]]]] * df.CD[[derivNames[[4]]]]
          }
          # Fraction of growth explained by Solow Residual
          solowFrac <- lambda_fd / df.CD[["dydtOvery"]]
          # Add to data frame
          df.CD$lambda_fd <- lambda_fd
          df.CD$solowFrac <- solowFrac
          # Add to the list of data frames
          dframes[[length(dframes)+1]] <- df.CD
        } else if (m == "CES"){
          if (e == "noE"){
            origCoeffs <- filter(AllCoeffs, model==m & energy==e & Country==Coun & flavor==flav)
            df.kl <- cbind(df, select(origCoeffs, -nest, -flavor, -energy))
            df.kl$nest <- "kl"
            n <- c(1,2)
            # Get the values for x1 and x2
            df.kl$x1 <- df.kl[[ nums[ unlist(n)[1]+1 ] ]] # GDP is first column, add 1 to get to the factors of production
            df.kl$x2 <- df.kl[[ nums[ unlist(n)[2]+1 ] ]]
            df.kl$dx1dt <- df.kl[[ derivNames[ unlist(n)[1]+1 ] ]]
            df.kl$dx2dt <- df.kl[[ derivNames[ unlist(n)[2]+1 ] ]]
            # Calculate derivative terms
            df.kl <- df.kl %>% 
              transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
              transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
              transform(lambda_fd = dydtOvery + 1/b * dbdtOverrho1) %>%
              transform(solowFrac = lambda_fd / dydtOvery)
            # Add to the list of data frames
            dframes[[length(dframes)+1]] <- df.kl
          } else if (e == "withE"){
            nests <- list(kle=c(1,2,3), lek=c(2,3,1), ekl=c(3,1,2))
            for (i in 1:3){
              n <- nests[i] # preserves name
              origCoeffs <- filter(AllCoeffs, model==m & energy==e & Country==Coun & flavor==flav & nest==names(n)[[1]])
              df.CESe <- cbind(df, select(origCoeffs, -nest, -flavor, -energy))
              df.CESe$nest <- names(n)[[1]]
              # Unpack names so that the formulae below are readable.
              df.CESe$x1 <- df.CESe[[ nums[ unlist(n)[1]+1 ] ]] # GDP is 1st column, add 1 to get factors of production
              df.CESe$x2 <- df.CESe[[ nums[ unlist(n)[2]+1 ] ]]
              df.CESe$x3 <- df.CESe[[ nums[ unlist(n)[3]+1 ] ]]
              df.CESe$dx1dt <- df.CESe[[ derivNames[ unlist(n)[1]+1 ] ]]
              df.CESe$dx2dt <- df.CESe[[ derivNames[ unlist(n)[2]+1 ] ]]
              df.CESe$dx3dt <- df.CESe[[ derivNames[ unlist(n)[3]+1 ] ]]
              # Calculate derivative terms
              df.CESe <- df.CESe %>% 
                transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
                transform(a = delta * b^(rho/rho_1) + (1 - delta) * x3^(-rho)) %>%
                transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
                transform(dadtOverrho = delta * b^(rho/rho_1 - 1) * dbdtOverrho1 - (1 - delta) * x3^(-(rho+1))*dx3dt) %>%
                transform(lambda_fd = dydtOvery + 1/a * dadtOverrho) %>%
                transform(solowFrac = lambda_fd / dydtOvery)
              # Add to the list of data frames
              dframes[[length(dframes)+1]] <- df.CESe
            }
          } else {
            stop (paste("Unknown e:", e))
          }
        } else {
          stop(paste("Unknown model", model))
        }
      }
    }
  }      
}
temp_data <- do.call(plyr::rbind.fill, dframes)
